---
title: "StatsBomb_Poland"
author: "Artur Czeszumski"
format: html
editor: visual
---

# Introduction

In this document I explore free StatsBomb data from World Cup 2022. I focus on the Polish national team. It is a first time I work with StatsBomb data. My goal is to explore the data and analyze the performance of the team.

# Libraries

Here we load all libraries required to get the data and execute the code in this file.

```{r}
library(tidyverse)
library(StatsBombR)
library(ggplot2)
library(SBpitch)
library(forcats)
```

# Select the data

Select World Cup 2022 (id 43) and events related to only Polish team.

This script can be used to analyze other teams too.

```{r}
team_to_analyze = "Poland" # specify a team
d <- FreeCompetitions() %>% # load all competitions 
  filter(competition_id == 43 & season_name == 2022) %>% # select World Cup 2022
  FreeMatches() %>% # load all matches
  filter(home_team.home_team_name == team_to_analyze | away_team.away_team_name == team_to_analyze) # select matchs of Polish team
```
